---
title: "02: Interactive Data Viz"
subtitle: "dyGraph Charts"
author: jmj00@ucsb.edu
output: 
  html_document:
    highlight: pygment
---

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      class.source = "numberLines lineAnchors",
                      warning = FALSE, message = FALSE,
                      eval = TRUE)
```

In the last [section](01_getting_data.html) we created two functions for getting the data for our COVID-19-Dashboard. This section will focus on building interactive charts. 

To experiement with these things, open a new test script. At the top of it source your helper code and load you data as covered in the last section:


```{r, echo = T}
source('./helpers.R')
covid19 = read_covid19('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv')
today   = today_centroids(counties, covid19)
```

## Charts

The dygraphs package is an R interface to the dygraphs JavaScript charting library. It provides rich utilities for visualizing time-series data in R. An exhastive list of functionalities and examples can be found [here](https://rstudio.github.io/dygraphs/).

Our goal for this section is to make a function, called `make_graph` that will take a FIP code as input and generate an interactive timeseries plot. The following code will get you a working example that you can customize and tweak using your knowledge from GEOG 183.

Let start by building our minimimally viable example:

```{r}
## Define FIP
FIP = 6037

# Filter the input data to the input FIP code
subset = filter(covid19, fips == FIP)

# DYGRAPHS uses the rownames of a data.frame to define the X-AXIS of a plot
# So lets set the rownames of the "subset" to the date attribute
rownames(subset) <- subset$date

# Select only the cases and deaths from the subset data and provide a title and axis labels
dygraph(data = select(subset, cases, deaths),
        main = paste0("COVID-19 Trend: ", subset$name[1]),
        ylab = 'Number of Cases/Deaths', 
        xlab = 'Date') 
```

Lets add some nice visual aspects using the `dyHighlight` option.

```{r}
dygraph(data = select(subset, cases, deaths),
        main = paste0("COVID-19 Trend: ", subset$name[1]),
        ylab = 'Number of Cases/Deaths', 
        xlab = 'Date') %>% 
# This options shades the area under the line plots
dyHighlight(highlightCircleSize = 4, 
            highlightSeriesBackgroundAlpha = 0.2,
            highlightSeriesOpts = list(strokeWidth = 2))
```

And the dyOptions features:

```{r}
dygraph(data = select(subset, cases, deaths),
        main = paste0("COVID-19 Trend: ", subset$name[1]),
        ylab = 'Number of Cases/Deaths', 
        xlab = 'Date') %>% 
# This options shades the area under the line plots
dyHighlight(highlightCircleSize = 4, 
            highlightSeriesBackgroundAlpha = 0.2,
            highlightSeriesOpts = list(strokeWidth = 2)) %>% 
dyOptions(stackedGraph = TRUE,
          colors = c("blue", "darkred"))
```

Lets add these thes fucntionalities to a function in our `helpers.R`. It might look something like this but should make use of your personal touches and aesthetics!

```{r}

make_graph  = function(covid19, FIP){

subset = filter(covid19, fips == FIP)
rownames(subset) <- subset$date

# Fit and exponetial model
exponential.model <- lm(log(cases)~ date, data = subset)
# use the model to predict a what a expoential curve would look like
subset$expCases = ceiling(exp(predict(exponential.model, list(date = subset$date))))

dygraph(data = select(subset, cases, deaths, expCases),
        main = paste0("COVID-19 Trend: ", subset$name[1]),
        ylab = 'Number of Cases/Deaths', 
        xlab = 'Date') %>% 
dyHighlight(highlightCircleSize = 4, 
            highlightSeriesBackgroundAlpha = 0.2,
            highlightSeriesOpts = list(strokeWidth = 2)) %>% 
dyOptions(colors = c("blue", "red", "black"))

}

```

To see it in action, lets call `make_graph` on FIP code 8031:

```{r}
make_graph(covid19, 8031)

```

## Conclusion

Fantastic, now we have a function for making a chart from our timeseries data. Lets move to the next [section](03_maps.html) where we will make interactive map of our centroid data using the `leaflet` package. 