---
title: "02: Interactive Data Viz"
subtitle: "Leaflet Maps"
author: jmj00@ucsb.edu
output: 
  html_document:
    highlight: pygment
---

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      class.source = "numberLines lineAnchors",
                      warning = FALSE, 
                      message = FALSE,
                      eval = TRUE,
                      out.width = "100%")
```


In the last [section](02_graphs.html) we created two functions for getting the data for our COVID-19-Dashboard. This section will focus on building interactive maps. To test these things, open a new test script. At the top of it attach the needed libraries and load your data as covered in the last section:

```{r, eval = TRUE}
source('./helpers.R')
covid19 = read_covid19('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv')
today   = today_centroids(counties, covid19)
```

Leaflet is one of the most popular open-source JavaScript libraries for interactive maps. It’s used by websites like The New York Times, The Washington Post,  GitHub and Flickr, as well as GIS specialists like OpenStreetMap, Mapbox, and CartoDB. The `leaflet` R package makes it easy to integrate and control Leaflet maps in R. An exhaustive list of functionality and examples can be found [here](https://rstudio.github.io/leaflet/).

Our goal for this section is to make a function, called `basemap` that will plot  county centroids with sizes based on there case count on a [tiled slippy map](https://en.wikipedia.org/wiki/Tiled_web_map).

We will create a second function - `zoom_to_county` - that zooms to a specific county based on a user supplied FIP code. The following code will get you a basic example that you can customize and tweak. 

## Web Tiles

Lets start by initialize a basic leaflet map. We start by giving our `today` data as the data input to the the generic `leaflet` call. We pass that map object to the `addProviderTiles` call. In this example we add the CartoDB.DarkMatter tileset but you can pick any tile set by name from the extensive leaflet provides found [here](https://leaflet-extras.github.io/leaflet-providers/preview/index.html). Finally we pass the map object with defined tiles to the `addScaleBar` function and ask for a scale bar in the bottom left of the map. The results look like:

```{r}
#Total Option:
length(providers)
```

```{r}
leaflet(data = today) %>%
# Add Tile sets from linked providers:
  addProviderTiles('CartoDB.DarkMatter') %>%
# Add a scale bar to a define position
  addScaleBar("bottomleft")

```

Note that our objects are not mapped! This is because we have not told `leaflet` how to represent spatial input To do so we will pass our map object to the `addCircleMarkers`:

```{r}
leaflet(data = today) %>%
  addProviderTiles(providers$CartoDB.DarkMatter) %>%
  addScaleBar("bottomleft") %>%
  addCircleMarkers()     
```

*Great!* They are now mapped but are pretty ugly. Lets make our map more pleasing by defining a color pallete and adding aestehtic parameters:

```{r}
# define a unique color pallete
pal <- colorNumeric("inferno", domain = today$size, n = 100)
  
map = leaflet(data = today) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addScaleBar("bottomleft") %>%
  addCircleMarkers(
    fillColor = ~pal(size), # the cicle inside color
    color = 'transparent',  # the circle border color
    fillOpacity = 0.5,      # The opacity of circle inside color
    radius = ~size*2,       # the circle size
    layerId = ~fips,        # the circle ID (needed for Shiny!)
    label   = ~name         # how should the markers be labeled?
  )

map
```


You job is to make this map your own by modifying the basemap,  color pallete, and design parameters in `addCircleMarkers.` The only thing that CANNOT be modified is the `layerID`. You will see why latter on.

## Map Modifiers

Here we are going to get a bit complicatd `r emo::ji("smile")`. What we want to do it define a function that will "zoom" to a defined county on your custom map. It other words, if a user asks about LA county, we want to zoom to LA county and show the county boundaries.

To do this we will make a new function that requires a leaflet map, county boudaries, and a FIP code.

```{r}

zoom_to_county = function(map, counties, FIP){
  # Filter the counties to the input FIP code
  shp = filter(counties, fips == FIP) 
  # Build a buffered bounding box to center the map on:
  bounds = shp %>% 
    # make bounding box
    st_bbox() %>% 
    # Make spatial
    st_as_sfc() %>% 
    # Buffer to .1 degree
    st_buffer(.1) %>% 
    # make new bounding box
    st_bbox() %>% 
    # extract coordinates as vector
    as.vector()
  # Clear all current shapes (remember county centroids are currently markers!)
    clearShapes(map) %>% 
  # Add the county shape making the outline color red and the fill an opaque white
    addPolygons(data = shp,
                color = "red",
                fillColor  = "white",
                fillOpacity = .2) %>% 
  # Fly the leaflet map to the buffered boundary
    flyToBounds(bounds[1], bounds[2], bounds[3], bounds[4])
}

```


Go ahead a copy the above code. Dont worry about changing any of it except for the way you want your Polygon to show up on the map. These are the parameters in the `addPolygon` method. 

Once completed, you can pass yourmap object to this new function to have you map zoom!

```{r}
zoom_to_county(map, counties, 6037)
```

Lets add these two functions to our `helpers.R`. They might look something like this but should make use of 
your personal touches and aesthetics!

```{r, eval = FALSE}

basemap = function(today){
  pal <- colorNumeric("inferno", domain = today$size, n = 50)
  pal2 <- colorNumeric("inferno", domain = today$cases, n = 50)
  
  leaflet(data = today) %>%
    addProviderTiles('CartoDB.Positron') %>%
    addScaleBar("bottomleft") %>%
    addCircleMarkers(
      fillColor   = ~pal(size), 
      color       = 'transparent', 
      fillOpacity = 0.5,
      radius      = ~size*2,
      layerId     = ~fips,
      label       = ~name ) %>% 
      addLegend("bottomright", pal = pal2, values = ~cases,
    title = paste("COVID Cases\n", max(today$date)),
    opacity = 1)
}

zoom_to_county = function(map, counties, FIP){
  
  shp = filter(counties, fips == FIP) 
  bounds = shp %>% 
    st_bbox() %>% 
    st_as_sfc() %>% 
    st_buffer(.1) %>% 
    st_bbox() %>% 
    as.vector()
  
  clearShapes(map) %>% 
    addPolygons(data = shp,
                color = "red", 
                fillColor  = "white", 
                fillOpacity = .2) %>% 
    flyToBounds(bounds[1], bounds[2], bounds[3], bounds[4])
}
```

# Testing 

```{r}
basemap(today)
```

## Conclusion

Fantastic, now we have a function for making a map from our centroid data. Lets move to the next [section](04_tables.html) where we will make interactive table of web-scraped Wikipedia data using the `DT` package. 
