---
title: "Autofill Table and Keyboard Listener"
output: 
  html_document:
    highlight: pygment
---

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      class.source = "numberLines lineAnchors",
                      warning = FALSE, message = FALSE,
                      eval = FALSE)
```

The last [section](08_shiny_tables.Rmd) we finalized the base Shiny Application. In this section we will add an advanced capability. This will be an auto complete text-box that allows users to select county names that are abel to modify the global FIP variable and change the interactive components.

# Extending the UI:

First lets add a text box to our UI. Now, we don't want to make users guess what their options are, so, we will add an autofilling capability that shows valid selections as a user types.

To do this, we need to load the `dqshiny` library (taken care of in the `helpers.R` script) and update the UI to include the autocomplete functionality. The way we set this up, the `autocomplete_input` function stores an object called `auto` in the `input` list. The default value of `auto` is `""` and the prompt displayed to users is "*Search for a County:*". 

```{r}
ui <- fluidPage(
  
  titlePanel('Mike Johnson: COVID-19 Tracker'),
  # Sidebar layout output definitions ----
  sidebarPanel(
    autocomplete_input("auto", "Search for a County:", 
                       value = "",
                       structure(today$fips, names = today$name)),
    # Output: Message ----
    textOutput("covid_message", container = h3),
    DTOutput("covid_table")
  ),
  
  # Main panel for displaying outputs ----
  mainPanel(
    # App title ----
    # Output: Map ----
    leafletOutput('covidmap'),
    # Output: Chart ----
    dygraphOutput('covid_chart')
  )
)
```

The tricky part is that we want to return a FIP code from the textbox (to parameterize `zoom_to_county`, `make_table`, and `make_chart`) but users need to be able to search by county/state names (bcause who really knows their FIP code ?!). Becasue of this, we need to structure the searchable data so that selecting a name, will return the associated FIP code. 

As you see above, we do this with the `structure` call where the first input (`today$fips`) defines what is returned to `auto` while the name input (`today$name`) defines the options users can search and select.

## Extending the Server

In the server function we are going to add an **observer**. This time we are not observing an event, rather an input object. The *expression* we will observe is to see if `auto` is either `""` (the default) or `NULL`. If it is, then the our reactive observer will return NULL. 

Otherwise, the observer will initate the same workflow we defined in our mouseclick event observer. Specifically,the global FIP is overridden, the covidmap zoom is updated, the covid_chart is re-generated, and a Wikipedianew table is visualized.

```{r}

observe( 
    if(input$auto == ""){ 
      NULL 
    } else {
      FIP <<- input$auto
      leafletProxy("covidmap") %>% zoom_to_county(counties, FIP)
      output$covid_chart <- renderDygraph({ make_graph(covid19, FIP) })
      output$covid_table      <- renderDT({ make_table(today, FIP) })
    }
  )
```


## Conclusion

Fantastic, now we have a completly functional and enhanced web application. Lets move to the next [section](10_publish.html) where we will apply a unique theme to our application and publish it as a website!