---
title: "02: Interactive Data Viz"
subtitle: "Tables"
author: jmj00@ucsb.edu
output: 
  html_document:
    highlight: pygment
---

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      class.source = "numberLines lineAnchors",
                      warning = FALSE, message = FALSE,
                      eval = TRUE)
```


```{r, echo = FALSE}
library(dplyr)
library(sf)   
library(DT)       
source('../helpers.R')
counties = readRDS("../data/counties.rds")
covid19  = read_covid19('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv')
today    = today_centroids(counties, covid19)
```


```{r, eval = FALSE}
## Load our helper file with our functions
source('./helpers.R')

## Load the county data
counties = readRDS("./data/counties.rds")

## Use our functions! from section 1:
covid19 = read_covid19('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv')
today = today_centroids(counties, covid19)
```

The `DT` package provides an R interface to the JavaScript library DataTables. R data objects (matrices or data frames) can be displayed as tables on HTML pages, and DataTables provides filtering, pagination, sorting, and many other features in the tables. Documentation and examples for DT can be found [here](https://rstudio.github.io/DT/)

The objective of this section is to build a function called `make_table` that takes a FIP code and build a nice summary table of information. We are going to look at two options. The first, will return a table of all counties in the same state as the selected FIP code. The second is a bit more complicated (and optional!) but will scape Wikipedia to get information about the county. Lets take a look at these options!

## Option 1: Sorting a Table by state & counts

```{r}
#Intial FIPS
FIP = 6037
# Fitler today's data to the FIP
myfip  = filter(today, fips == FIP)

# Filter todays data to the state of the input FIP
mydata = filter(today, state == myfip$state) %>% 
# Arrange the cases from largest to smallest
  arrange(desc(cases)) %>% 
# Drop the geometry column
  st_drop_geometry() %>% 
# Keep the  County name, cases and deaths
  select(County = county, Cases = cases, Deaths = deaths) %>% 
# Create a new variable called Death Rate
  mutate(DeathRate = paste0(100* round(Deaths/Cases,2), "%")) %>% 
  head(20)


# Make an interactive Table! with a caption
datatable(mydata, caption = paste('COVID-19 Statistics', myfip$state, myfip$date), options = list(paging = FALSE, searching = FALSE)) 
```

## Option 2: Scraping Wikipedia Data

```{r}
library(rvest)

# Filter the input data to the input FIP code
county_here = filter(today, fips == FIP)
# Build a Wikipedia URL using the data attribute 'name'
url = paste0('https://en.wikipedia.org/wiki/',  gsub(" ", "_", county_here$name))
  # Steps:
  ## 1. Read the HTML from the URL
  ## 2. Read the Nodes from the HTML page and select the table with a class infobox
  ## 3. Convert raw HTML node to a data.frame
links <- read_html(url) %>%
           html_nodes("table.infobox") %>% 
           html_table(fill= TRUE) 


## Make a datatable from the resulting data.frame and turn off paging
datatable(links[[1]],  caption = paste('Wikipedia Information:', county_here$name), 
              options = list(paging = FALSE, searching = FALSE, ordering = FALSE))
```


Lets add all of these to the helper.R file which looks something like:


```{r}

make_table = function(today, FIPS){
  county_here = filter(today, fips == FIP)

  url = paste0('https://en.wikipedia.org/wiki/',  gsub(" ", "_", county_here$name)) %>% 
           read_html(url) %>%
           html_nodes("table.infobox") %>% 
           html_table(fill= TRUE) 

  l = links[[1]]
  ll = l[!l[,1] == l[,2],]

## Make a datatable from the resulting data.frame and turn off paging
  datatable(ll,  caption = paste('Wikipedia Information:', county_here$name), 
              options = list(paging = FALSE, searching = FALSE, ordering = FALSE))
}

```

A complete `helpers.R` file can be found [here](https://github.com/mikejohnson51/COVID-19-dashboard/blob/master/helpers.R) for download / copy-paste. 

Notice that all the visualization functions we have built rely on our `county`, `timeseries`, or `today` data and a usr defined `FIP code`. From these we can make informative plots, maps and tables. 

```{r, eval = FALSE}
# Base Example
map = basemap(today)

zoom_to_county(map, couties, 6037)

make_table(today, 6037)

make_graph(covid19, 6037)
```

The purpose of our Shiny Application will be to display these resoruces on a common page, and allow user input (from sources like clicks and text-input) to change the FIP code and the resulting visualizations. To do this we need a speacial type of programing known as 'reative'.

